Does It Scale? Symbolic Execution of PHP Web Applications

### 2017-01-23 paper.tex
5,6,7 refactored

\begin{document}
\title{Does It Scale? Symbolic Execution of PHP Web Applications}
%\titlenote{Produces the permission block, and copyright information}
\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

%\author{Stefan Mühlbauer}
\affiliation{%
  %\institution{Technische Universität Braunschweig}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

%\author{Christian Kästner}
%\orcid{1234-5678-9012}
\affiliation{%
  %\institution{Carnegie Mellon University}
  %\streetaddress{5000 Forbes Avenue}
  %\city{Pittsburgh} 
  %\state{PA} 
  %\postcode{15217}
}
%\email{s.muehlbauer@tu-bs.de}

%\author{Tien N. Nguyen}
%\orcid{1234-5678-9012}
\affiliation{%
  %\institution{University of Texas at Dallas}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

@@

\begin{abstract}
Dynamic web applications have become widely popular and are to a large portion
based on the scripting language PHP. Dynamic languages like PHP challenge
static analysis, as they generate code at runtime. One approach to unfold their
staged nature is symbolic execution by approximating client page output.
Knowledge about possible output enables a range of developer tools and analyses
for vulnerability detection.

However, previous tools have only been evaluated for smaller systems and
scalability for larger systems has not been shown. We criticize the missing
evalaution of practicality of respective tools and ask whether symbolic
execution for output approximation for PHP is scalable and what limitations can
be faced. This paper presents our experiences with re-implementing and
extending a previous symbolic execution engine for PHP that we evaluate for
several large PHP systems.
We find that dynamic language features are likely to be resolved imprecisely
for symbolic values and impede symbolic execution. We explore
strategies to mitigate limitations based on previous work on dynamic symbolic
execution by manually providing concrete rather than symbolic values.
\end{abstract}

@@

A common characteristic of all dynamic web applications is their
\emph{staged nature}: A dynamic web application as a whole
consists of both static code, such as scripts, and dynamically generated code,
such as client page output. The latter code, although assembled at
runtime, may contain client-side parts of the web application, such as
JavaScript. So, to study dynamic web applications in its entirety, we need to
consider both static as well as dynamic aspects of systems.

@@

To investigate the question, whether we can have a practical and scalable
symbolic executoin engine, we re-implemented the engine with the specifications
of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
as object-oriented programming. We evaluate our
symbolic execution engine for several large-scale and modern PHP systems.

@@

Based on observations we identify conceptual limitations of symbolic
execution for PHP.  Dynamic features such as indirect function invocations and
include expression require concrete information to be evaluated properly. Since
these expressions may be assembled at runtime and can contain symbolic
information, for many cases this restricts a symbolic execution engine from
further execution.
Although some effort to statically approximate include expression can be spent,
it is often non-trivial since expressions can contain information coming from
various inputs including databases, user inputs or configuration files. We
empirically evaluate our observations for a corpus including large, actively maintained PHP systems.

@@

\end{lstlisting}
		\caption{
			Approximated client page output of the server-side code in Figure
			\ref{fig:web_application}. The directives in lines 6 and 8 highlight output
			that can be repeated an arbitrary number of times, the directives in line 10
			and 14 mark two alternative output parts.
			\label{fig:output_example}
		}
    \end{subfigure}
    \vspace{2mm}
    \caption{
    	Illustration of static ouptut approximation using symbolic
    	execution
    	\label{fig:code_example}
    }
\end{figure}

Figure \ref{fig:code_example} shows two code listings. The snippet of server-side web
application code in Figure \ref{fig:web_application} contains both HTML code
(input form in lines 1 -- 4) as well as PHP code: First, the function
\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
times. Second, the function is indirectly called using the built-in mechnism
\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
\texttt{4} as arguments.  Finally, the input from the form in lines 1 – 4 is
used: The variable \texttt{\$greeting} consists of three concatenated strings,
\texttt{Hello}, the value of input field \texttt{name}, and an exclamation mark.
For the greeting there are two possible client page output
variants: If the input field \texttt{name} is empty at runtime, line 19 is
executed and \texttt{No name entered!} is printed, or, if a name was entered,
line 21 is executed and the greeting constructed in line 17 is printed.

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	ouput approximation a callgraph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.

@@

\subsubsection{Symbolic/Concrete Loop Execution}
Throughout the implementation we basically followed the symbolic execution
semantics of \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928}. As described
in Section \ref{sec:workflow} for loops, only one iteration is executed since
the number of iterations is indeterminable for most cases. While this is the
case for while and do-while loops, the number of loop iterations for for or
foreach loops is generally determinable and known before entering the loop
body. For loops use a loop counter that usually is incremented or decremented
after each loop iteration until a loop condition is not satisfied anymore;
foreach loops simply iterate over a given array value.
Our execution semantics follows this scheme and simulates concrete execution
for those types of loops if possible. Otherwise, for instance, if a loop
counter, loop counter limit for for loops, or variable for foreach loops is
symbolic, we resort to executing only a single iteration. This design choice
was made since it is more precise and, in contrast to only considering one
iteration in all cases, slightly increased code coverage for \sf{WordPress}.

@@

\subsubsection{Dynamic Language Features}
\label{sec:experiment_dynamicfeatures} 
Aside from the previous issues, we identified two dynamic language features
that we were not able to execute accurately. Dynamic features in PHP are
features that take as input rather a expression that is resolved at runtime
than a static constant value. For \sf{WordPress}, we faced missing files and
functions that are included using dynamic includes and dynamic function
invocation respectively.

First, file inclusion in PHP usually is managed by statements like \tt{include},
\tt{include\_once} or \tt{require} and \tt{require\_once}. For transitive
includes, statements ending with \tt{\_once} will not include the same file
twice. The include statements make an attempt to include a file, but do not
return a warning in contrast to the require statements. Each of these
statements takes as input either a string constant (static include) or an
expression that is evaluated at runtime (dynamic include). The string that the
expression evaluates to represents an absolute or relative paths to the target
file that the interpreter finally attempts to include.
Figure \ref{fig:dynamic_includes} shows a small example of a dynamic include containing
information retrieved from a database. Since any input for a symbolically
executed application is symbolic this information can be propagated to include
expressions. Also, configuration files represent only one particular
configuration or even just default values or placeholders. Hence, symbolic
execuion is prone to fail for dynamic includes if symbolic information is contained.

Second, function can either be invocated directly by statically providing the
name of the function (or method), or indirectly. Indirect, or dynamic function
invocation in PHP are enabled by using built-in functions like \tt{call\_user\_func},
\tt{call\_user\_func\_array} or simply passing a string value with a direct
function call to the \tt{eval} function. The first two functions take as
arguments the name of the function two call, and all arguments to that
functions either as additional arguments or a single associative array
respectively.
The \tt{eval} function evaluates a given PHP expression passed as a string
literal.
Again, we encountered issues similar to dynamic include expressions with
\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
\tt{header\_sserif} which both print a HTML headline. Depending on whether the
value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
\tt{sserif}, a different function is called. Given that the function name is 
concrete at runtime, the desired function can be called. Otherwise, if the
function name is assembled at runtime and contains symbolic values, no function
can be called as the name is ambiguous or unknown. For this example, symbolic
execution is not able to determine which function to call indirectly.

@@

\section{Evaluation} \label{sec:evaluation}
As already stated in the previous Section \ref{sec:experience_report}, we
encountered several problems during the implementation of our symbolic
execution engine for \sf{WordPress}. The main items of critique of the symbolic
execution engine and tools described in previous work
\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
are the missing support for widely-used language features, and the evaluation
only using small-scale systems. Although evaluating a tool for a small systems
can be a valuable proof-of-concept, it does not necessarily cover the question
of whether a tool is practical on a larger-scale. Therefore in the following we
present our methods to evaluate practicality of our symbolic output
approximation. Moreover, we investigate to which extent the conceptual limitations encountered
with \sf{WordPress} apply for a wider choice of PHP systems.

\subsection{Experiment Setup} \label{sec:experiment_setup}
We have encountered several issues for \sf{WordPress} that conceptually limit
the scalability of symbolic execution. Since our observations so far only considers
one example system an, we evaluate our symbolic execution engine for a wider
range of PHP systems with regard to the following questions. The metrics to
approach those questions are explained in Sections
\ref{HowAccurateIsOurApproximation} and \ref{sec:understanding_limitations}.

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{PhpSync}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928}
has shown high code coverage for small systems. As our symbolic execution
re-implements \sf{PhpSync} and extends with support for additional language
features, we ask: Can  we replicate high coverage results at least on a
small-scale. Therefore, our PHP corpus contains four small-scale systems used
in previous work \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask: Do system size and usage of modern language features
factors affect the practicality of our symbolic execution engine? We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which conceptual
limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study addressing the feature usage in PHP
systems. Some more detailed description of this case study can be found in
Section \ref{sec:related_work}. For the small-scale systems we selected five systems from
a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of system than we intended
to since the parser used in our implementation,
\sf{Quercus}\footnote{\sf{Quercus} is an open-source Java implementation of PHP
5 by Caucho Technology. We reused the parser front-end for our implementation.
\sf{Quercus} can be obtained from \url{http://quercus.caucho.com}.}, did not
support a number of language features.
The full list of PHP systems for that we evaluate our symbolic execution engine along with descriptive code metrics can be found in Table \ref{corpus}.

@@

\subsection{Measuring Approximation Success} \label{heuristic}
In order to accurately answer the question how good our symbolic execution
engine approximates client page output, we would require knowledge of all
possible variants to compare our approximation against. Since we do not have
any ground truth information alike, we approach this demand by using a
heuristic. Rather than having knowledge about all client page output variants
we say our approximation is accurate if all string literals (which are embedded
in and  scattered across server-side code) that may eventually become part of any
client page output variant (in the following referred to as \emph{output
candidate}) is contained in our symbolic output approximation.
It is non-trivial to determine whether a string literal is an output candidate,
so we heuristically classify string literals as output candidates if they
contain the characters \tt{<}, \tt{>}, or both since we expect output to contain
HTML tags.

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 96 percent of the string literals are classified correctly. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually were responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but could not do so without decreasing precision. 
In spite of missing half of the output candidates, we decided
to use a simple, yet precise and distinctive classifier. We rather work on a
smaller sample subset of actual output candidates with less false positives
than a larger set of string literals, where more are mistakenly classified as
output candidates.

\subsection{Measuring Approximation Accuracy}
\label{HowAccurateIsOurApproximation} 
In Section \ref{heuristic} we introduced the definition of output candidates as
expected output. For an approximation to be most accurate, it needs to contain
all output candidates of a system analyzed. Based on whether output
candidates are reached by an execution path and part of the output
approximatiopn define two metrics to approximate measuring accuracy.

@@

The coverage results for our case study are illustrated in Figure
\ref{coverage}. We could replicate high coverage results for the four small-scale systems with a reach
coverage and output coverage over 80 percent respectively. For the three
modern/large-scale systems we measured poor coverage with reach coverage
ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
We measured medium to high coverage for the more recent small-scale systems:
reach coverage ranging from 43 to 85 percent, output coverage ranging from 40
to 85 percent.

\subsection{Understanding Limitations} \label{sec:understanding_limitations}
As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
coverage metrics were poor for large-scale and small-scale/recent systems. To
understand what output candidates we missed and why, we further investigated
our approximation results and conducted three more measurements.

@@

	\end{subfigure}
	\caption{Evaluation results for output candidates, include expressions and
	callgraph analysis for not-reached output candidates}
\end{figure}

\subsubsection{Where did we miss output?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out
whether their surrounding program code was not accessed/accessible, and if so,
why. We are interested in those output candidates that are located in HTML
files (possibly with nested PHP), or located in a function definition. Plain
HTML files do not require any further execution and represent output candidates
that just need to be included properly to be part of the output. Aside, output
candidates that are part of a function definition are only missed if the
corresponding function is never called at runtime.

We started classifying missed output candidates by their string literal
context. As the classification statistics in Figure \ref{coverage} illustrate,
for almost all systems tested (except for \textsf{AddressBook} and
\textsf{TimeClock}) not-reached output candidates had a function context.

@@

For a function to be never reached there are several scenarios: A function is
undefined at runtime if the corresponding script file is never included. In
turn, if the function is defined at runtime, the function is dead
code if there is no call site for this function, or the function is only called
indirectly and the function name could not be resolved accurately. In PHP there
are several ways to call a function beside direct call sites. 
%The language offers indirect call mechanisms like callback-commands or an
% evaluation function that parses and evaluates PHP source code as strings.

Given a symbolic value, indirect function invocation mechanisms like
callback-commands are likely to fail since target function name and the symbolic value do not match.
This also applies for the evaluation of include expressions as concrete values
representing include targets can be included, but any include target containing
symbolic values is ambiguous. Since it is unlikely that all functions containing
missed output candidates are dead code having no call site (direct or
indirect) at all, in Section \ref{WhyDidFunctionCallsFail} we further
investigate how many functions fail due to inaccurately resolved indirect
function invocations. Moreover, we evaluate the accuracy of evaluation of
include expressions in the next section.

\subsubsection{(Dynamic) Includes}
\label{WhyDidIncludesFail}
From Section \ref{sec:experiment_dynamicfeatures} we learn that include
expressions can not be evaluated accurately for symbolic values, and from the
previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed output
candidates can be explained by simply missing to include a file. To better
understand the impact of include expression resolution throughout analyzing
systems, we construct two additional metrics. First, we are interested in how
many include expressions throughout the execution we actually reach in a system.
Therefore we introduce the metric \emph{reach coverage for include expressions}
as the ratio of reached include expressions and the total number of reach
expressions in a system to check whether failed includes may have caused
subsequent include expressions and so on. Second, we measure how successful
include expressions are resolved.

Figure \ref{fig:include_coverage_results} illustrates the classification of
include expressions for each system: We can either reach an include expression
and successfully resolve it, or not. Or, we did not reach an include expression at all. 
Figure \ref{fig:include_coverage_results} indicates that except for the
small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not really
reach a great portion of include expressions, although for most systems the
resolution success rate was over 80 percent. Note that one failed attempt
to include a script file can result in a cascade of missing more include
expressions as the include target script file is not included and executed.

\subsubsection{Dynamic Function Invocations}
\label{WhyDidFunctionCallsFail}
Aside from failed or successful includes, this metric describes one scenario of
why functions containing not-reached output candidates are never called: The
trivial explanation would be that there are no direct calls of those functions,
yet PHP offers a number of ways to call a function indirectly, for instance
using built-in functions, or simply evaluating strings as PHP code using the
\texttt{eval} function.

Moreover we need to take into account that one missed function execution can
result in missing even more function calls (direct or indirect) and executions.
Thus, for functions containing not-reached output candidates we measure whether
they can be (1) only, (2) partially or (3) never accessed transitively from
functions that have no direct call sites. We refer to those functions (or
access points) having no direct call site as callback candidates since the only
way to reach them is through dynamic function invocation mechanisms. Measuring
how many not-reached output candidates can be explained by functions that are callback
candidates helps to understand whether, and if so, indirect function invocation
mechanisms and their usage may impact our analysis’ code coverage.

>>>>

\begin{document}
\title{Does It Scale? Symbolic Execution of PHP Web Applications}
%\titlenote{Produces the permission block, and copyright information}
%\subtitle{Output-oriented Symbolic Execution for Static Output Approximation}
%\subtitlenote{The full version of the author's guide is available as  \texttt{acmart.pdf} document}

\author{Stefan Mühlbauer}
\affiliation{%
  \institution{Technische Universität Braunschweig}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  \state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Christian Kästner}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{Carnegie Mellon University}
  %\streetaddress{5000 Forbes Avenue}
  %\city{Pittsburgh} 
  %\state{PA} 
  %\postcode{15217}
}
%\email{s.muehlbauer@tu-bs.de}

\author{Tien N. Nguyen}
%\orcid{1234-5678-9012}
\affiliation{%
  \institution{University of Texas at Dallas}
  %\streetaddress{P.O. Box 1212}
  %\city{Braunschweig} 
  %\state{Germany} 
  %\postcode{43017-6221}
}
%\email{s.muehlbauer@tu-bs.de}

@@

\begin{abstract}
Dynamic web applications have become widely popular and are to a large portion
based on the scripting language PHP. Dynamic languages like PHP challenge
static analysis, as they generate code at run-time. One approach to unfold their
staged nature is symbolic execution by approximating client page output.
Knowledge about possible output enables a range of developer tools and analyses
for vulnerability detection. However, previous tools have only been evaluated
for smaller systems and scalability for larger systems has not been shown.

In this paper, we criticize the limited expressiveness of the evaluation of
practicality for respective tools. We ask whether symbolic execution for output
approximation for PHP is scalable, and what limitations can be faced. By
extending a symbolic execution engine based on previous work and evaluating it
for some larger PHP systems we find dynamic features to be likely to hinder
symbolic execution. We explore strategies to mitigate limitations inspired by
work on dynamic symbolic execution and aim to push for more comprehensive
evaluation for developer tools.
\end{abstract}

@@

A common characteristic of all dynamic web applications is their
\emph{staged nature}: A dynamic web application as a whole
consists of both static code, such as scripts, and dynamically generated code,
such as client page output. The latter code, although assembled at
run-time, may contain client-side parts of the web application, such as
JavaScript. So, to study dynamic web applications in its entirety, we need to
consider both static as well as dynamic aspects of systems.

@@

To investigate the question, whether we can have a practical and scalable
symbolic execution engine, we re-implemented the engine with the specifications
of the previous symbolic execution semantics \cite{Nguyen:2014:BCG:2635868.2635928} for PHP and additional features, such
as object-oriented programming. We evaluate our
symbolic execution engine for several large-scale and modern PHP systems.

@@

Based on observations we identify conceptual limitations of symbolic
execution for PHP.  Dynamic features such as indirect function invocations and
include expression require concrete information to be evaluated properly. Since
these expressions may be assembled at run-time and can contain symbolic
information, for many cases this restricts a symbolic execution engine from
further execution.
Although some effort to statically approximate include expression can be spent,
it is often non-trivial since expressions can contain information coming from
various inputs including databases, user inputs or configuration files. We
empirically evaluate our observations for a corpus including large, actively maintained PHP systems.

@@

\end{lstlisting}
		\caption{
			Approximated client page output of the server-side code in Figure
			\ref{fig:web_application}. The directives in lines 6 and 8 highlight output
			that can be repeated an arbitrary number of times, the directives in line 10
			and 14 mark two alternative output parts.
			\label{fig:output_example}
		}
    \end{subfigure}
    \vspace{2mm}
    \caption{
    	Illustration of static output approximation using symbolic
    	execution
    	\label{fig:code_example}
    }
\end{figure}

Figure \ref{fig:code_example} shows two code listings. The snippet of server-side web
application code in Figure \ref{fig:web_application} contains both HTML code
(input form in lines 1 -- 4) as well as PHP code: First, the function
\texttt{make\_titles} defined in lines 7 – 13 prints the string \texttt{\$arg} passed as an argument between zero and an arbitrary number of
times. Second, the function is indirectly called using the built-in mechanism
\texttt{call\_user\_func} in line 15 with the string \texttt{Headline} and
\texttt{4} as arguments.  Finally, the input from the form in lines 1 – 4 is
used: The variable \texttt{\$greeting} consists of three concatenated strings,
\texttt{Hello}, the value of input field \texttt{name}, and an exclamation mark.
For the greeting there are two possible client page output
variants: If the input field \texttt{name} is empty at run-time, line 19 is
executed and \texttt{No name entered!} is printed, or, if a name was entered,
line 21 is executed and the greeting constructed in line 17 is printed.

@@

	\item \emph{Varis}: Editor services such as “jump to declaration” are
	nontrivial for dynamic web applications due to their staged nature. Varis
	provides editor services across stages for client-side code: Starting with an
	output approximation a call-graph with conditional edges is constructed, which
	allows navigation in client-site code (HTML, JS and CSS) although it is
	embedded in server-side code
	\cite{Nguyen:2015:VIS:2819009.2819140,Nguyen:2014:BCG:2635868.2635928}.

@@

\subsubsection{Symbolic/Concrete Loop Execution}
Throughout the implementation we basically followed the symbolic execution
semantics of \sf{PhpSync} \cite{Nguyen:2014:BCG:2635868.2635928}. As described
in Section \ref{sec:workflow} for loops, only one iteration is executed since
the number of iterations is indeterminable for most cases. While this is the
case for while and do-while loops, the number of loop iterations for for or
for-each loops is generally determinable and known before entering the loop
body. For loops use a loop counter that usually is incremented or decremented
after each loop iteration until a loop condition is not satisfied anymore;
for-each loops simply iterate over a given array value.
Our execution semantics follows this scheme and simulates concrete execution
for those types of loops if possible. Otherwise, for instance, if a loop
counter, loop counter limit for for loops, or variable for for-each loops is
symbolic, we resort to executing only a single iteration. This design choice
was made since it is more precise and, in contrast to only considering one
iteration in all cases, slightly increased code coverage for \sf{WordPress}.

@@

\subsubsection{Dynamic Language Features}
\label{sec:experiment_dynamicfeatures} 
Aside from the previous issues, we identified two dynamic language features
that we were not able to execute accurately. Dynamic features in PHP are
features that take as input rather a expression that is resolved at run-time
than a static constant value. For \sf{WordPress}, we faced missing files and
functions that are included using dynamic includes and dynamic function
invocation respectively.

First, file inclusion in PHP usually is managed by statements like \tt{include},
\tt{include\_once} or \tt{require} and \tt{require\_once}. For transitive
includes, statements ending with \tt{\_once} will not include the same file
twice. The include statements make an attempt to include a file, but do not
return a warning in contrast to the require statements. Each of these
statements takes as input either a string constant (static include) or an
expression that is evaluated at run-time (dynamic include). The string that the
expression evaluates to represents an absolute or relative paths to the target
file that the interpreter finally attempts to include.
Figure \ref{fig:dynamic_includes} shows a small example of a dynamic include containing
information retrieved from a database. Since any input for a symbolically
executed application is symbolic this information can be propagated to include
expressions. Also, configuration files represent only one particular
configuration or even just default values or placeholders. Hence, symbolic
execution is prone to fail for dynamic includes if symbolic information is contained.

Second, function can either be invocated directly by statically providing the
name of the function (or method), or indirectly. Indirect, or dynamic function
invocation in PHP are enabled by using built-in functions like \tt{call\_user\_func},
\tt{call\_user\_func\_array} or simply passing a string value with a direct
function call to the \tt{eval} function. The first two functions take as
arguments the name of the function two call, and all arguments to that
functions either as additional arguments or a single associative array
respectively.
The \tt{eval} function evaluates a given PHP expression passed as a string
literal.
Again, we encountered issues similar to dynamic include expressions with
\ref{fig:indirect_calls} illustrates two functions \tt{header\_serif} and
\tt{header\_sserif} which both print a HTML headline. Depending on whether the
value of the variable \tt{\$style} retrieved from a database is \tt{serif} or
\tt{sserif}, a different function is called. Given that the function name is 
concrete at run-time, the desired function can be called. Otherwise, if the
function name is assembled at run-time and contains symbolic values, no function
can be called as the name is ambiguous or unknown. For this example, symbolic
execution is not able to determine which function to call indirectly.

@@

\section{Evaluation} \label{sec:evaluation}
As already stated in the previous Section \ref{sec:experience_report}, we
encountered several problems during the implementation of our symbolic
execution engine for \sf{WordPress}. The main items of critique of the symbolic
execution engine and tools described in previous work
\cite{Nguyen:2011:AFH:2190078.2190142,Nguyen:2014:BCG:2635868.2635928,Nguyen:2015:CPS:2786805.2786872,Nguyen:2015:VIS:2819009.2819140}
are the missing support for widely-used language features, and the evaluation
only using small-scale systems. Although evaluating a tool for a small systems
can be a valuable proof-of-concept, it does not necessarily cover the question
of whether a tool is practical on a larger-scale. Therefore in the following we
present our methods to evaluate practicality of our symbolic output
approximation. Moreover, we investigate to which extent the limitations encountered
with \sf{WordPress} apply for a wider choice of larger PHP systems.

\subsection{Experiment Setup} \label{sec:experiment_setup}
We have encountered several problems for \sf{WordPress} that conceptually limit
the scalability of symbolic execution. Since our observations so far only consider
one example system an, we evaluate our symbolic execution engine for a wider
range of PHP systems with regard to the following questions. The metrics to
approach those questions are explained in Sections
\ref{HowAccurateIsOurApproximation} and \ref{sec:understanding_limitations}.

First, the evaluation of the output approximation using the symbolic execution
engine, \sf{PhpSync}, in previous work \cite{Nguyen:2014:BCG:2635868.2635928}
has shown high code coverage for small systems. As our symbolic execution
re-implements \sf{PhpSync} and extends with support for additional language
features, we ask: \textit{Can  we replicate high coverage results at least on a
small-scale?} Therefore, our PHP corpus contains four of the small-scale systems
that \sf{PhpSync} has been evaluated for \cite{Nguyen:2014:BCG:2635868.2635928}.

Second, we ask: \textit{Do system size and usage of modern language features
factors affect the practicality of our symbolic execution engine?} We
selected both a range of large-scale and small-scale modern PHP systems to
better separate and understand the circumstances under which limitations might affect practicality. For the selection of large-scale systems
we borrow three systems from a case study by Hills et al. \cite{hills2015variable} addressing the
feature usage patterns and frequencies in PHP systems. Some more detailed description of this
case study can be found in Section \ref{sec:related_work}. For our choice of modern small-scale systems we
selected five systems from a list of recent Content Management Systems (CMS) written in PHP \cite{codegeekz}. For both
selections we were bound to select a smaller number of systems than we intended
to since the parser used in our implementation,
\sf{Quercus}\footnote{\sf{Quercus} is an open-source Java implementation of PHP
5 by Caucho Technology. We reused the parser front-end for our implementation.
\sf{Quercus} can be obtained from \url{http://quercus.caucho.com}.}, did not
support a number of language features.
The full list of PHP systems for that we evaluate our symbolic execution engine along with descriptive code metrics can be found in Table \ref{corpus}.

@@

\subsection{Measuring Approximation Success} \label{heuristic}
In order to accurately answer the question how good our symbolic execution
engine approximates client page output, we would require knowledge of all
possible variants to compare our approximation against. Since we do not have
any ground truth information alike, we approach this demand by using a
heuristic. Rather than having knowledge about all client page output variants
we say our approximation is accurate if all string literals (which are embedded
in and  scattered across server-side code) that may eventually become part of any
client page output variant is contained in our symbolic output approximation. In
the following we refer to those expected output string literals as \emph{output candidates}.

With respect to the staged nature of web applications it is
non-trivialstatically to determine whether a string literal is an output
candidate, so we heuristically classify string literals as output candidates if
they contain the characters \tt{<}, \tt{>}, or both since we expect output to
contain HTML tags.

We evaluated this heuristic manually with a sample of 400 string
literals randomly selected from our corpus of PHP systems (see Table
\ref{corpus}). We measured for our heuristic classifier a precision of 94 percent, and a
recall of 50 percent. This means that six percent of the string literals are
classified incorrectly as false positives. In turn, the classifier is highly
distinctive as 94 percent of the output candidates are classified as such. The
recall of 50 percent means that half of the string literals, which we manually 
classified as output candidates, actually are responsive to the classifier. We
attempted to increase recall by looking for further distinctive properties to build a
classifier from, but were not able to do so without decreasing precision. 
In spite of missing half of the output candidates, we decided
to use a simple, yet precise and distinctive classifier. We rather work on a
smaller sample subset of actual output candidates with less false positives
than a larger set of string literals, where more are mistakenly classified as
output candidates.

\subsection{Measuring Approximation Accuracy}
\label{HowAccurateIsOurApproximation} 
In Section \ref{heuristic} we introduced the definition of output candidates as
expected output. For an approximation to be accurate, it needs to contain
all output candidates of a system analyzed. Based on whether output
candidates are reached by an execution path and part of the output
approximation we define two metrics to approximate measuring accuracy.

@@

The coverage results for our case study are illustrated in Figure
\ref{coverage}. For our first evaluation question in Section \ref{sec:experiment_setup} we
could replicate high coverage results for the four small-scale systems with a reach
coverage and output coverage over 80 percent respectively. Whereas for the
three modern/large-scale systems we measured poor coverage with reach coverage
ranging from 5 to 30 percent and output coverage ranging from 4 to 20 percent.
We measured medium to high coverage for the more recent small-scale systems:
reach coverage ranging from 43 to 85 percent, output coverage ranging from 40
to 85 percent.

\subsection{Understanding Limitations} \label{sec:understanding_limitations}
As we have seen in Section \ref{HowAccurateIsOurApproximation} both measured
coverage metrics were poor for large-scale and small-scale/recent systems. To
understand what output candidates we missed and why, we further investigate
our approximation results and conduct three subsequent measurements.

@@

	\end{subfigure}
	\caption{Evaluation results for output candidates, include expressions and
	call-graph analysis for not-reached output candidates}
\end{figure}

\subsubsection{Where did we miss output?}\label{WhatLiteralsDidWeMiss?}
Our initial approach to understand missed output candidates is to find out
whether their surrounding program code cannot not accessed, and if
so, why. We are interested in those output candidates that are located in HTML
files (possibly with nested PHP), or located in a function definition. Plain
HTML files do not require any further execution and represent output candidates
that just need to be included properly to be part of the output. Aside, output
candidates that are part of a function definition are only missed if the
corresponding function is never invocated at run-time, or simply dead code.

We started classifying missed output candidates by their string literal
context. As the classification statistics in Figure \ref{coverage} illustrate,
for almost all systems tested (except for \textsf{AddressBook} and
\textsf{TimeClock}) not-reached output candidates have a function context.

@@

For a function to never be reached there are several scenarios: A function is
undefined at run-time if the corresponding script file is never included. In
turn, if the function is defined at run-time, the function is dead
code if there is no call site for this function, or the function is only called
dynamically and the function name could not be resolved precisely. In PHP there
are several ways to call a function beside static call sites, as described in
Section \ref{sec:experiment_dynamicfeatures}.
%The language offers indirect call mechanisms like callback-commands or an
% evaluation function that parses and evaluates PHP source code as strings.

Given a symbolic value, indirect function invocation mechanisms like
callback-commands are likely to fail since target function name and the symbolic
value the expression evaluates to do not match.
This also applies for the evaluation of include expressions as concrete values
representing include targets can be included, but any include target containing
symbolic values is ambiguous. Since it is unlikely that all functions containing
missed output candidates are dead code having no call site (static or dynamic)
at all, in Section \ref{WhyDidFunctionCallsFail} we further investigate how
many functions fail due to inaccurately resolved indirect function invocations.
Moreover, we evaluate the accuracy of evaluation of include expressions in the next section.

\subsubsection{(Dynamic) Includes}
\label{WhyDidIncludesFail}
From Section \ref{sec:experiment_dynamicfeatures} we learn that include
expressions containic symbolic valuescannot be resolution precisely, and from
the previous Section \ref{WhatLiteralsDidWeMiss?} that very little of missed
output candidates can be explained by simply missing to include a file. To
better understand the impact of include expression resolution throughout
analyzing systems, we construct two additional metrics.

First, we are interested in how many include expressions throughout the
execution we actually reach in a system. Therefore we measure the number of
reached include expressions in a system to check whether failed includes may
have caused subsequent include expressions and so on.

Second, we measure how many reached include expressions are
succesfully resolved throughout the analysis for each system.

Figure \ref{fig:include_coverage_results} illustrates the classification of
include expressions for each system: We can either reach an include expression
and successfully resolve it, or not. Or, we did not reach an include expression at all. 
Figure \ref{fig:include_coverage_results} indicates that except for the
small-scale systems as well as \sf{Anchor} and \sf{Nibbleblog} we did not
reach a great portion of include expressions, although for most systems the
resolution success rate (ratio of resolved and reached include expressions) was
over 80 percent. Note that one failed attempt to include a script file can 
result in a cascade of missing more include expressions as the include target
script file is not included and executed. This may explain the low number of
reached include expressions, but relatively high success rate.

\subsubsection{Dynamic Function Invocations}
\label{WhyDidFunctionCallsFail}
Aside from failed or successful includes, this metric describes one scenario of
why functions containing not-reached output candidates are never invocated: The
trivial explanation would be that there are no static invocations of those
functions, yet PHP offers a number of ways to invocate a function dynamically,
for instance using built-in functions, or simply evaluating strings as PHP code
at run-time using the \texttt{eval} function.

Moreover we need to take into account that one missed function execution can
result in missing even more function invocations.
Thus, for functions containing not-reached output candidates we measure whether
they can be (1) only, (2) partially or (3) never accessed transitively from
functions that have no direct call sites. We refer to those functions having no
static call site as \emph{callback candidates} since the only way to reach them
is through dynamic function invocation mechanisms. Measuring how many not-reached output candidates can be explained by functions that are callback
candidates helps to understand whether, and if so, indirect function invocation
mechanisms and their usage may impact our analysis’ code coverage.
